Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2024-04-20 revision-1/hyper.tex
checkpoint

@@

\begin{theorem}
  Gradual release implies passive security.
\end{theorem}

@@

\begin{theorem}
  Robust declassification implies malicious security.
\end{theorem}

>>>>

\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{X}{Y}$ iff for all
    $\store \in \mems(X \cup Y)$ we have
  $\margd{\pmf}{X \cup Y}(\store) =
  \pmf(\store_X) * \pmf(\store_Y)$
\end{definition}

@@

@@

\begin{theorem}
  Passive security with robust declassification implies malicious security.
\end{theorem}

### 2024-04-20 revision-1/lang.tex
checkpoint

We let $x$ range over \emph{variables}  which are identifiers
where client ownership is specified- e.g., $\elab{\mesg{\mathit{foo}}}{\cid}$
is a message $\mathit{foo}$ that was sent to $\cid$. We let $X$
range over sets of variables, and more specifically, $S$ ranges over sets of secret variables $\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables $\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables $\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables $\rvl{w}$, and $O$ ranges over sets of output variabels $\out{\cid}$.
Given a program $\prog$, we write $\iov(\prog)$ to
denote the set of $S \cup M \cup P \cup O$ of variables in $\prog$
with ownership made explicit, and we write $\flips(\prog)$ to
denote the set $R$ of random samplings in $\prog$ with ownership
made explicit. We write
$\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
of variables $X$ and parties $P$, we write $X_P$ to denote the subset
of $X$ owned by any party in $P$, in particular we write $X_H$ and $X_C$ to
denote the subsets belonging to honest and corrupt parties,
respectively.

@@

First, clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus
\flip{a}$ and $\secret{y} \fplus \flip{b}$ where $\secret{x}$ and
$\secret{y}$ are 1 and 2's secrets respectively.  Then they both
perform a secure opening of $\secret{x} \fplus \flip{a}$ and
$\secret{y} \fplus \flip{b}$, by authenticating each others' shares,
and record them as $\mesg{d}$ and $\mesg{e}$ respectively. Note
that the secrets remain hidden since $\flip{a}$ and $\flip{b}$
are random values.
$$
\begin{array}{lcl}
  \elab{\mesg{a}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{1}\\
  \elab{\mesg{a}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{2}\\
  \elab{\mesg{b}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{1}\\
  \elab{\mesg{b}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{2}\\
  \elab{\mesg{d}}{1} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
  \elab{\mesg{e}}{1}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
  \elab{\mesg{d}}{2} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
  \elab{\mesg{e}}{2}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
\end{array}
$$
Now, each client uses their share of $\secret{c}$ to generate their
share of the final output, which is again reported via a secure opening.
Homomorphic key calculations are elided for brevity.
$$
\begin{array}{lcl}
  \rvl{1} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{1}\\
  \rvl{2} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{2}\\
  \out{1} &:=& \elab{(\macopen\ \rvl{1} \ \macauth(\rvl{2},\ldots))}{1} \\
  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \rvl{2})}{2} 
\end{array}
$$
Some information about input secrets is revealed by the product, but
secure opening ensures high integrity and hence robustness to
corruption, aka robust declassification.

>>>>

We let $x$ range over \emph{variables} which are identifiers where
client ownership is specified- e.g.,
$\elab{\mesg{\mathit{foo}}}{\cid}$ is a message $\mathit{foo}$ that
was sent to $\cid$. We let $X$ range over sets of variables, and more
specifically, $S$ ranges over sets of secret variables
$\elab{\secret{w}}{\cid}$, $R$ ranges over sets of random variables
$\elab{\flip{w}}{\cid}$, $M$ ranges over sets of message variables
$\elab{\mesg{w}}{\cid}$, $P$ ranges over sets of public variables
$\rvl{w}$, and $O$ ranges over sets of output variables $\out{\cid}$.
Given a program $\prog$, we write $\iov(\prog)$ to denote the set $S
\cup M \cup P \cup O$ of variables in $\prog$ with ownership made
explicit and $\secrets(\prog)$ to denote $S$, and we write
$\flips(\prog)$ to denote the set $R$ of random samplings in $\prog$
with ownership made explicit. We write $\vars(\prog)$ to denote
$\iov(\prog) \cup \flips(\prog)$. For any set of variables $X$ and
parties $P$, we write $X_P$ to denote the subset of $X$ owned by any
party in $P$, in particular we write $X_H$ and $X_C$ to denote the
subsets belonging to honest and corrupt parties, respectively.

@@

First, clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus
\flip{a}$ and $\secret{y} \fplus \flip{b}$ where $\secret{x}$ and
$\secret{y}$ are 1 and 2's secrets respectively.  Then they both
perform a secure opening of $\secret{x} \fplus \flip{a}$ and
$\secret{y} \fplus \flip{b}$, by authenticating each others' shares,
and record them as $\mesg{d}$ and $\mesg{e}$ respectively. Note
that the secrets remain hidden since $\flip{a}$ and $\flip{b}$
are random values.
$$
\begin{array}{lcl}
  \elab{\mesg{a}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{1}\\
  \elab{\mesg{a}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{2}\\
  \elab{\mesg{b}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{1}\\
  \elab{\mesg{b}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{2}\\
  \elab{\mesg{d}}{1} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{} \fminus \mack{\macx{\flip{a}}{\Oracle}}{}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
  \elab{\mesg{e}}{1}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{} \fminus \mack{\macx{\flip{b}}{\Oracle}}{}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
  \elab{\mesg{d}}{2} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{} \fminus \mack{\macx{\flip{a}}{\Oracle}}{}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
  \elab{\mesg{e}}{2}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{} \fminus \mack{\macx{\flip{b}}{\Oracle}}{}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
\end{array}
$$
Now, each client uses their share of $\secret{c}$ to generate their
share of the final output, which is again reported via a secure opening.
Homomorphic key calculations are elided for brevity.
$$
\begin{array}{lcl}
  \rvl{1} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{1}\\
  \rvl{2} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{2}\\
  \out{1} &:=& \elab{(\macopen\ \rvl{1} \ \macauth(\rvl{2},\ldots))}{1} \\
  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \rvl{2})}{2} 
\end{array}
$$
Some information about input secrets is revealed by the product, but
secure opening ensures high integrity and hence robustness to
corruption, aka robust declassification.

### 2024-04-20 revision-1/model.tex
checkpoint

\section{Security Model}

@@

The simulator is represented by a probabilistic algorithm $\SIM_C$,
aka a \emph{simulation}, that is parameterized by corrupt inputs and
the output of an ideal functionality, and that returns a set of
adversarial views (as a memory) with some probability. Given
corrupt inputs $\store$ and ideal functionality output $v$,  
we write
$
\prob(\SIM(\store,v) = \store')
$
to denote the probability that $\SIM(\store,v)$
returns corrupt views $\store'$ as a result. We can then define the
probability distribution of corrupt views reconstructed
by the simulator as follows:
\begin{definition}
  Given $C$, $\store$, and $v$, we write $\dist(\SIM(\store,v))$ to
  denote the distribution of corrupt views reconstructed by the
  simulation, where for
  all $\store' \in \mems(V)$:
  $$
  \dist(\SIM(\store,v))(\store')\ \defeq\ \prob(\SIM(\store,v) = \store') 
  $$
\end{definition}

Then we can define passive security in the real/ideal
model as follows. 
\begin{definition}[Passive Security]
  Assume given a program $\prog$ that correctly implements an ideal
  functionality $\idealf$, with $\iov(\prog) = (S,M,O)$.  Then $\prog$
  is \emph{passive secure in the simulator model} iff for all
  partitions of the federation into honest and corrupt sets $H$ and $C$
  with $|C| < |H|$ and for all $\store \in \mems(S)$ there exists a
  simulation $\SIM$ such that:
  $$
  \dist(\SIM(\store_{S_C},\idealf(\store))) = \condd{\progtt(\prog)}{M_C}{\store}
  $$
\end{definition}

@@

\begin{definition}[Corrupt Inputs, Honest Outputs]
  Given a program $\prog$ with $\iov(\prog) = (S,M,O)$ , define $\cinputs$ as the
  messages in $M$ sent from corrupt to honest parties:
  $$
  \cinputs = \{\ \elab{\mesg{w}}{\cid}\ \mid\  \elab{\mesg{w}}{\cid} \in M \wedge \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
  \wedge \cid \in H \wedge \cid' \in C \ \} 
  $$
  and similarly define $\houtputs$ as the messages in $M$ sent from honest to corrupt parties.
  %Define also $(\afilter\ \prog)$ as $\prog$ with all instructions of the form $\eassign{\mesg{w}}{\cid}{\be}{\cid'}$ removed
  %where $\cid \in H \wedge \cid' \in C$.
\end{definition}

>>>>

\section{Security Model}
\label{section-pmf}

\emph{Simulator security} in the MPC setting means that protocols do
not reveal any secret information other than what is implicitly
declassified by the output of the ideal functionality that the
protocol implements. Characterization of both real world protocol
execution and simulation is defined probabilistically. Following
previous work \cite{barthe2019probabilistic} we use probability mass
functions to express joint dependencies between input and output
variables, as a metric of information leakage.

\subsection{Probability Mass Functions}

We define joint probability mass functions (pmfs) in the standard
manner, though following \cite{barthe2019probabilistic} we use
memories to denote mappings of variables to values (i.e., outcomes),
so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
program containing these variables.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ for some $X$ to values in $\mathbb{F}_p$, such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ the \emph{marginal distribution} of variables $X$
  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
  $$
  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
  $$
\end{definition}

\begin{definition}
  Given $\margd{\pmf}{X}$, let $\stores$ be a set of memories with the
  same domain $Y \subseteq X$. Then the \emph{conditional distribution given
  $\stores$}  denoted
  $\condd{\pmf}{X}{\stores}$ is a distribution with domain $X$ where for all
  $\store \in \mems(X)$:
  $$
  \condd{\pmf}{X}{\stores}(\store) =
  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
  $$
  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
\end{definition}
To recover familiar notation we allow the syntactic
sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
%\begin{eqnarray*}
%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
%\end{eqnarray*}

\subsection{Basic Distribution of a Protocol}
Now we can define the probability distribution of a program $\prog$,
that we denote $\progtt(\prog)$. Since $\fedcat$ is deterministic the
results of any run are determined by the input secrets together with
the random tape. And since we constrain programs to not overwrite
views, we are assured that \emph{final} memories contain both a
complete record of all initial secrets as well as views resulting from
communicated information.

We consider random tapes to be selected from a uniform distribution
modulo relevant pre-processing, e.g., Beaver triples and MACed share
distributions.  We will also assume that any possible initial secrets
are equally likely when formulating the program distribution.
We define $\runs(\prog)$ as the set of final memories resulting
from execution of $\prog$ given any initial memory, and
treat all elements of $\runs(\prog)$ as equally likely.
This establishes the basic program distribution that can be
marginalized and conditioned to quantify input/output
information dependencies. 
%In this
%setting, given a program $\prog$ with $\iov(\prog) = S \cup V$ and
%$\flips(\prog) = F$ we will consider all $\store \in \mems(S \cup V
%\cup F)$ such that $ \config{\store_{S \cup F}}{\prog} \redxs
%\config{\store_}{\varnothing} $ to be equally probable, establishing
%the basic distribution of the program. %From this, we can immediately
%derive the marginal distribution of $S \cup V$ to reason about
%dependencies between secrets and views.
\begin{definition}
  \label{def-progd}
  \label{definition-progd}
  Given program $\prog$ with $\secrets(\prog) = S$ and $\flips(\prog) = R$, and
  any assumed pre-processing predicate $\preproc$ on memories, define:
  $$
  \runs(\prog) \defeq \{ \store \mid \preproc(\store) \wedge
  (\dom(\store) = \iov(\prog) \cup R) \wedge (\store_{S \cup R},\prog) \redxs (\store,\varnothing) \}
  $$
  Then the \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
  $$
  \progtt(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$
  In some cases, we will also be concerned with the (joint)
  probabilities of expression interpretation given a preceding program
  execution, and we write $\progtt(\prog, \be)$ to denote the program
  distribution $\progtt(\prog;\itv := \be)$ where $\itv$ is a
  special variable that is never used in programs.
\end{definition}

@@

Information about honest secrets can be revealed to corrupt clients
through messages sent from honest to corrupt clients, and through
publicly broadcast information from honest clients. Dually,
corrupt clients can impact protocol integrity through the messages
sent from corrupt to honest clients, and through publicly broadcast information
from corrupt clients. We call the former \emph{corrupt views}, and
the latter \emph{honest views}. Generally we let $V$ range over sets
of views.
\begin{definition}[Corrupt and Honest Views]
  Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as the messages in
  $V = \views(\prog)$ sent from honest
  to corrupt parties, called \emph{corrupt views}:
  $$
  \houtputs = \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog
  \wedge \cid \in H \ \} 
  \cup \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
  \wedge \cid \in C \wedge \cid' \in H \ \} 
  $$
  and similarly define $\cinputs$ as the messages in $V$ sent from corrupt to honest
  parties, called \emph{honest views}:
  $$
  \cinputs = \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog
  \wedge \cid \in C \ \} 
  \cup \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in \prog
  \wedge \cid \in H \wedge \cid' \in C \ \} 
  $$
\end{definition}

In the passive setting the simulator must construct a probabilistic
algorithm $\SIM$, aka a \emph{simulation}, that is parameterized by
corrupt inputs and the output of an ideal functionality, and that
returns a distribution of corrupt views that is indistinguishable
from the distribution of corrupt views in the real world
protocol execution.

Given inputs $\store$ and ideal
functionality output $v$, we write $ \prob(\SIM(\store,v) = \store') $
to denote the probability that $\SIM(\store,v)$ returns corrupt views
$\store'$ as a result. We can then define the probability distribution
of corrupt views reconstructed by the simulator as follows:
\begin{definition}
  Given $C$, $\store$, and $v$, we write $\dist(\SIM(\store,v))$ to
  denote the distribution of corrupt views reconstructed by the
  simulation, where for
  all $\store' \in \mems(V)$:
  $$
  \dist(\SIM(\store,v))(\store')\ \defeq\ \prob(\SIM(\store,v) = \store') 
  $$
\end{definition}
Then we can define passive security in the real/ideal
model as follows. 
\begin{definition}[Passive Security]
  Assume given a program $\prog$ that correctly implements an ideal
  functionality $\idealf$, with $\views(\prog) = V$.  Then $\prog$
  is \emph{passive secure in the simulator model} iff for all
  partitions of the federation into honest and corrupt sets $H$ and $C$
  and for all $\store \in \mems(S)$ there exists a
  simulation $\SIM$ such that:
  $$
  \dist(\SIM(\store_{C},\idealf(\store))) = \condd{\progtt(\prog)}{\houtputs}{\store}
  $$
\end{definition}

@@
